---
title: "first_script"
author: "merve ozgul"
date: "1/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


*Analysis on World Happiness Dataset *

In this project I will analyze World Happiness dataset, which has 155 countries and different attributes. I took the dataset from Kaggle: https://www.kaggle.com/unsdsn/world-happiness/home



#1) Loading the Dataset & Initial Exploration of Dataset

I use "Import Dataset" at the Global Environment to load the dataset. Then I convert the dataset to dataframe. 

```{r}
#library(readr)
#X2015 <- read_csv("~/Desktop/world-happiness-report/2015.csv")
happy_2015 <- as.data.frame(X2015)
```

I use some functions to have a glance at the dataset and understand it. I would like to check the column(attribute) names initially, to see if I should make any changes before starting to further explore the dataset. I like to change the names, optimize them in a meaningful and shorter manner. This way it makes it easier for me to call column names, in the future.
```{r}
#Viewing column names
names(happy_2015)
```
It would be better to change all column names in a way that they won't include white space in their names. It is also good to note their original name in the metadata, in case the attribute name we create is not clear enough.

 [1] "Country"                       "Region"                       
 [3] "Happiness Rank"                "Happiness Score"              
 [5] "Standard Error"                "Economy (GDP per Capita)"     
 [7] "Family"                        "Health (Life Expectancy)"     
 [9] "Freedom"                       "Trust (Government Corruption)"
[11] "Generosity"                    "Dystopia Residual"   

```{r}
#Changing column names
names(happy_2015) <- c("country", "region", "happiness_rank", "happiness_score", "std_error", "gdp_per_cpt", "family", "life_exp", "freedom", "trust_corruption", "generosity", "dystopia_residual")
```

#Characteristics of the Dataset
Now we can start with our analysis: 

```{r}
head(happy_2015)

#Viewing stats about each attribute
summary(happy_2015)

#structure
str(happy_2015)

#missing values 
sum(is.na(happy_2015))

```

There is no missing values in our dataset, which will accelerate our analysis. We don't have to put any time to handle the missing values.

```{r}
#Number of countries in our dataset
length(unique(happy_2015$country))

#Number of regions
length(unique(happy_2015$region))
```

Since we have 10 regions and 158 countries, we can subset our dataset to regions and look at the distributions. 

#Subsetting our dataset
```{r}
unique(happy_2015$region)

#Subsetting Western Europe
w_europe <- happy_2015[which(happy_2015$region == "Western Europe"), ]

#North America
n_america <- happy_2015[which(happy_2015$region == "North America"), ]

#Latin America & Caribbean
n_america <- happy_2015[which(happy_2015$region == "Latin America and Caribbean"), ]

#Middle East and Northern Africa
m_east_n_africa <- happy_2015[which(happy_2015$region == "Middle East and Northern Africa"), ]

#Australia & New Zealand
aust_newzealand <- happy_2015[which(happy_2015$region == "Australia and New Zealand"), ]

#Central and Eastern Europe
central_easteu <- happy_2015[which(happy_2015$region == "Central and Eastern Europe"), ]

#Eastern Asia
east_asia <- happy_2015[which(happy_2015$region == "Eastern Asia"), ]

#



```














